Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-11-07 text/ms.tex
Catch up on edits to the text
Bringing in figure changes and captions
Switch fitted to fit

To obtain estimates of the probability and magnitude of black swans, we fitted population dynamic models to abundance time series from around the world. For each population, we estimated the shape of the process noise tails by measuring the degrees of freedom ($\nu$) of a Student t-distribution.

@@

Our main analysis focuses on the commonly applied Gompertz population dynamics model \citep[e.g.][]{knape2012,dennis2014,connors2014}. The Gompertz model represents population growth as a linear function in $\log$ space. If we let $x_t$ represent the $\log$ abundance ($N$) at time $t$, we can represent the Gompertz model as:
\begin{align*}
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma).
\end{align*}
The growth parameter $\lambda$ represents the expected growth rate if $N_t = 1$. The model is density independent if $b = 1$, maximally density dependent if $b = 0$, and inversely density dependent if $b < 0$. Usually, the process noise $\epsilon_t$ is modelled as normally distributed, but in our paper we assume it is drawn from a t distribution with scale parameter $\sigma$ and degrees of freedom $\nu$. In previous analyses of the GPDD, the Gompertz is most often identified as the most parsimonious population model that can be fitted to these data \citep{brook2006}.

@@

We fitted all models in a Bayesian framework using Stan \citep{stan-manual2014} via the R computing environment \citep{r2014}. Stan samples from the posterior distribution with an adaptive version of Hamiltonian Markov chain Monte Carlo called the No-U-Turn Sampler and generally obtains less correlated samples than algorithms such as the Gibbs sampler \citep{hoffman2014}. We tested to ensure that the chains had sufficiently converged and that the sampler had obtained sufficient independent samples from the posterior ($\widehat{R} < 1.05$, $n_\mathrm{eff} > 200$; Supporting Information).

@@

% TODO fitted vs. fit
We fitted alternative population models to test if four key phenomena systematically changed our conclusions. Autocorrelation has been suggested as a reason for increased observed variability of abundance time series through time, which could create apparent heavy tails \citep{inchausti2002}; therefore, we fitted a model that included serial correlation in the residuals. Additionally, previous work has modelled abundance or growth rates without accounting for density dependence \citep{halley2002,segura2013}; therefore, we fitted a simpler model in which we assumed density independence. Third, observation error could bias parameter estimates \citep{knape2012} or mask our ability to detect heavy tails \citep{ward2007}; therefore, we fit a model where we allowed for a fixed quantity of observation error ($0.2$ standard deviations on a log scale). Finally, the Gompertz model assumes that population growth rate declines linearly with log abundance. We also fit an alternative model, the Ricker-logistic model, which assumes that population growth rate declines linearly with abundance itself (Supporting Information).

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg.pdf}
\caption{
An illustration of fitting population dynamic models that allow for heavy tails, represented by the Student-t degrees of freedom parameter $\nu$. (a, b) The probability density for t distributions with a scale parameter of 1 and different values of $\nu$. Small values of $\nu$ create heavy tails. As $\nu$ approaches infinity the distribution approaches the normal distribution. For example, at $\nu = 2$, the probability of drawing a value more than five standard deviations below the mean is 1.8\%, whereas the probability of drawing such a value from a normal distribution is nearly zero ($2.9\cdot10^{-5}$\%). (c--e) Simulated population dynamics from a Gompertz model with process noise drawn from t distributions with three different values of $\nu$. Coloured dots in panels c and d represent jumps with less than a 1 in 1000 chance of occurring in a normal distribution. (f--h) Estimates of $\nu$ from models fit to the times series in panels c--e. Shown are the posterior samples (histograms), median and interquartile range of the posterior (IQR) (dots and line segments), and the exponential prior on $\nu$ (dashed lines). Colour shading behind panels f--h illustrates the region of heavy tails.
}
\label{fig:didactic}
\end{center}
\end{figure}

>>>>

To obtain estimates of the probability and magnitude of black swans, we fit population dynamic models to abundance time series from around the world. For each population, we estimated the shape of the process noise tails by measuring the degrees of freedom ($\nu$) of a Student t-distribution.

@@

Our main analysis focuses on the commonly applied Gompertz population dynamics model \citep[e.g.][]{knape2012,dennis2014,connors2014}. The Gompertz model represents population growth as a linear function in $\log$ space. If we let $x_t$ represent the $\log$ abundance ($N$) at time $t$, we can represent the Gompertz model as:
\begin{align*}
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma).
\end{align*}
The growth parameter $\lambda$ represents the expected growth rate if $N_t = 1$. The model is density independent if $b = 1$, maximally density dependent if $b = 0$, and inversely density dependent if $b < 0$. Usually, the process noise $\epsilon_t$ is modelled as normally distributed, but in our paper we assume it is drawn from a t distribution with scale parameter $\sigma$ and degrees of freedom $\nu$. In previous analyses of the GPDD, the Gompertz is most often identified as the most parsimonious population model that can be fit to these data \citep{brook2006}.

@@

We fit all models in a Bayesian framework using Stan \citep{stan-manual2014} via the R computing environment \citep{r2014}. Stan samples from the posterior distribution with an adaptive version of Hamiltonian Markov chain Monte Carlo called the No-U-Turn Sampler and generally obtains less correlated samples than algorithms such as the Gibbs sampler \citep{hoffman2014}. We tested to ensure that the chains had sufficiently converged and that the sampler had obtained sufficient independent samples from the posterior ($\widehat{R} < 1.05$, $n_\mathrm{eff} > 200$; Supporting Information).

@@

We fit alternative population models to test if four key phenomena systematically changed our conclusions. Autocorrelation has been suggested as a reason for increased observed variability of abundance time series through time, which could create apparent heavy tails \citep{inchausti2002}; therefore, we fit a model that included serial correlation in the residuals. Additionally, previous work has modelled abundance or growth rates without accounting for density dependence \citep{halley2002,segura2013}; therefore, we fit a simpler model in which we assumed density independence. Third, observation error could bias parameter estimates \citep{knape2012} or mask our ability to detect heavy tails \citep{ward2007}; therefore, we fit a model where we allowed for a fixed quantity of observation error ($0.2$ standard deviations on a log scale). Finally, the Gompertz model assumes that population growth rate declines linearly with log abundance. We also fit an alternative model, the Ricker-logistic model, which assumes that population growth rate declines linearly with abundance itself (Supporting Information).

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg2.pdf}
\caption{
An illustration of fitting population dynamic models that allow for heavy tails, represented by the Student-t degrees of freedom parameter $\nu$. (a, b) The probability density for t distributions with a scale parameter of 1 and different values of $\nu$. Small values of $\nu$ create heavy tails. As $\nu$ approaches infinity the distribution approaches the normal distribution. For example, at $\nu = 2$, the probability of drawing a value more than five standard deviations below the mean is 1.8\%, whereas the probability of drawing such a value from a normal distribution is nearly zero ($2.9\cdot10^{-5}$\%). (c--e) Simulated population dynamics from a Gompertz model with process noise drawn from t distributions with three different values of $\nu$. Coloured dots in panels c and d represent jumps with less than a 1 in 1000 chance of occurring in a normal distribution. (f--h) Estimates of $\nu$ from models fit to the times series in panels c--e. Shown are the posterior samples (histograms), median and interquartile range of the posterior (IQR) (dots and line segments), and the exponential prior on $\nu$ (dashed lines). Colour shading behind panels f--h illustrates the region of heavy tails.
}
\label{fig:didactic}
\end{center}
\end{figure}

### 2014-11-07 text/supp.tex
Switch fitted to fit

\item We filled in all missing time steps with \texttt{NA} values and imputed single missing values with the geometric mean of the previous and following values. We chose a geometric mean to be linear on the log scale that the Gompertz and Ricker-logistic models were fitted on.

\item We filled in single recorded values of zero with the lowest non-zero value in the time series \citep[following][]{brook2006a}. This assumes that single values of zero result from abundance being low enough that censusing overlooked individuals that were actually present. We turned multiple zero values in a row into \texttt{NA} values. This implies that multiple zero values were either censusing errors or caused by emigration. Regardless, our population models were fitted on a multiplicative (log) scale and so could not account for zero abundance. To avoid distorting the original data too strongly, we removed populations in which we filled in more than four zeros.

@@

We fitted our models with Stan 2.4.0 \citep{stan-manual2014}, and R 3.1.1 \citep{r2014}. We began with four chains and $2000$ iterations, discarding the first $1000$ as warm up (i.e.~4000 total samples). If $\hat{R}$ (the potential scale reduction factor --- a measure of chain convergence) was greater than $1.05$ for any parameter or the minimum effective sample size, $n_\mathrm{eff}$, (a measure of the effective number of uncorrelated samples) for any parameter was less than $200$, we doubled both the total iterations and warm up period and sampled from the model again. These thresholds are in excess of the minimums recommended by \citet{gelman2006a} of $\hat{R} < 1.1$ and effective sample size $> 100$ for reliable point estimates and confidence intervals. In the majority of cases our minimum thresholds were greatly exceeded. We continued this procedure up to $8000$ iterations ($16000$ total samples) by which all chains were deemed to have sufficiently converged. These chain lengths may seem low to those familiar with software such as WinBUGS or JAGS, but the No-U-Turn Hamiltonian Markov chain Monte Carlo Sampler in Stan generally requires far fewer iterations to obtain equivalent effective sample sizes \citep{stan-manual2014}.

@@

We considered a version of the Gompertz model in which an autoregressive
parameter was fitted to the process noise residuals:
\begin{align*}
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, \phi \epsilon_{t-1}, \sigma).
\end{align*}
In addition to the parameters in the original Gompertz model, we estimate an additional parameter $\phi$, which represents the dependence of subsequent residuals. Based on the results of previous analyses with the GPDD \citep[e.g.][]{connors2014} and the chosen priors in previous analyses \citep[e.g.][]{thorson2014a} and to greatly speed up chain convergence when running our model across all populations, we placed a weakly informative prior on $\phi$ that assumed the greatest probability density near zero with the reduced possibility of $\phi$ being near $-1$ or $1$. Specifically, we chose $\phi \sim \mathrm{Truncated\mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. The MCMC chains did not converge for \modelsNoConvergeAROne\ populations according to our criteria ($\widehat{R} < 1.05, n_\mathrm{eff} > 200$) after 8000 iterations of four chains. This included only \modelsNoConvergeAROneHeavyBase\ populations in which Pr($\nu < 10$) $> 0.5$ categorized them as heavy in the main Gompertz model. We did not include these models in Fig.~\ref{fig:alt}. \textit{TODO perhaps I should go up to 16,000 so that more converge.}

@@

We also fitted a Ricker-logistic model:
\begin{align*}
x_t &= x_{t-1} + r_{\mathrm{max}}\left(1 - \frac{N_{t-1}}{K}\right) + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma),
\end{align*}
where $r_\mathrm{max}$ represents the theoretical maximum population growth rate that is obtained when $N_t$ (abundance at time $t$) $= 0$. The parameter $K$ represents the carrying capacity and, as before, $x_t$ represents the $\log$ transformed abundance at time $t$. The Ricker-logistic model assumes a linear decrease in population growth rate with increases in abundance. In contrast, the Gompertz model assumes a linear decrease in population growth rate with increases in \textit{log} abundance ($x_t$) \citep[e.g.][]{thibaut2012}.

@@

First, we tested the ability to estimate $\nu$ given different true values of $\nu$ and different sample sizes. We took stochastic draws from t distributions with different $\nu$ values ($\nu = 3, 5, 10,$ and $10^6$ [$\approx$ normal]), with central tendency parameters of $0$, and scale parameters of $1$. We started with $1600$ stochastic draws and then fitted the models again at the first $800, 400, 200, 100, 50,$ and $25$ draws. Each time we recorded the posterior samples of $\nu$.

@@

We then fitted our Gompertz models to the simulated datasets with all parameters (except $\nu$) set near the median values estimated in the GPDD. We repeated this with $50$ and $100$ samples without observation error, $50$ samples with observation error ($\sigma_\mathrm{obs} = 0.2$), and $50$ samples with the same observation error and a Gompertz model that allowed for correctly specified observation error magnitude.

@@

We fitted a multilevel beta regression model to the predicted probability of heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed dynamics. The beta distribution is useful when response data range on a continuous scale between zero and one. We used a logit link function as is typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$ represents the predicted value for population $i$, class $j$, order $k$, and species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$ represents a vector of predictors (such as lifespan) for population $i$ with associated $\beta$ coefficients. The intercepts are allowed to vary from the overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$), taxonomic order ($\alpha^\mathrm{order}_k$), and species ($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \; \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \; \mathrm{species}}$. Where possible, we also allowed for error distributions around the predictors by incorporating the standard deviation of the posterior samples for the Gompertz parameters $\lambda$, $b$, and $\log \sigma$ around the mean point value as normal distributions (not shown in the above equation).

@@

We incorporated weakly informative priors into our model: $\mathrm{Cauchy}(0,
10)$ on the global intercept $\alpha$, $\mathrm{Half\mhyphen Cauchy}(0, 2.5)$ on all standard deviation parameters, $\mathrm{Half\mhyphen Cauchy}(0, 10)$ on the dispersion parameter $\phi_\mathrm{disp}$, and $\mathrm{Cauchy}(0, 2.5)$ on all other parameters \citep{gelman2006c, gelman2008d}. Compared to normal priors, the Cauchy priors concentrate more probability density around expected parameter values while allowing for a higher probability density far into the tails, thereby allowing the data to dominate the posterior more strongly if it disagrees with the prior. Our conclusions were not qualitatively changed by using uniform priors. We fitted our models with 5000 total iterations per chain, 2500 warm-up iterations, four chains, and discarding every second sample to save memory. We checked for chain convergence visually and with the same criteria as before ($\widehat{R} < 1.05$ and $n_\mathrm{eff} >200$ for all parameters).

To derive taxonomic-order-level estimates of the probability of heavy tails accounting for time-series length (Fig \ref{fig:order-estimates}), we fitted a separate multilevel model with the same structure but with only $\log$ time-series length as a predictor. (In this case, we did not want to control for intrinsic population characteristics such as density dependence.) Since our predictors were centered by subtracting their mean value, we obtained order-level estimates of the probability of heavy tails at mean log time-series length by adding the posteriors for $\alpha$, $\alpha^\mathrm{class}_j$, and $\alpha^\mathrm{order}_k$.

>>>>

\item We filled in all missing time steps with \texttt{NA} values and imputed single missing values with the geometric mean of the previous and following values. We chose a geometric mean to be linear on the log scale that the Gompertz and Ricker-logistic models were fit on.

\item We filled in single recorded values of zero with the lowest non-zero value in the time series \citep[following][]{brook2006a}. This assumes that single values of zero result from abundance being low enough that censusing overlooked individuals that were actually present. We turned multiple zero values in a row into \texttt{NA} values. This implies that multiple zero values were either censusing errors or caused by emigration. Regardless, our population models were fit on a multiplicative (log) scale and so could not account for zero abundance. To avoid distorting the original data too strongly, we removed populations in which we filled in more than four zeros.

@@

We fit our models with Stan 2.4.0 \citep{stan-manual2014}, and R 3.1.1 \citep{r2014}. We began with four chains and $2000$ iterations, discarding the first $1000$ as warm up (i.e.~4000 total samples). If $\hat{R}$ (the potential scale reduction factor --- a measure of chain convergence) was greater than $1.05$ for any parameter or the minimum effective sample size, $n_\mathrm{eff}$, (a measure of the effective number of uncorrelated samples) for any parameter was less than $200$, we doubled both the total iterations and warm up period and sampled from the model again. These thresholds are in excess of the minimums recommended by \citet{gelman2006a} of $\hat{R} < 1.1$ and effective sample size $> 100$ for reliable point estimates and confidence intervals. In the majority of cases our minimum thresholds were greatly exceeded. We continued this procedure up to $8000$ iterations ($16000$ total samples) by which all chains were deemed to have sufficiently converged. These chain lengths may seem low to those familiar with software such as WinBUGS or JAGS, but the No-U-Turn Hamiltonian Markov chain Monte Carlo Sampler in Stan generally requires far fewer iterations to obtain equivalent effective sample sizes \citep{stan-manual2014}.

@@

We considered a version of the Gompertz model in which an autoregressive
parameter was fit to the process noise residuals:
\begin{align*}
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, \phi \epsilon_{t-1}, \sigma).
\end{align*}
In addition to the parameters in the original Gompertz model, we estimate an additional parameter $\phi$, which represents the dependence of subsequent residuals. Based on the results of previous analyses with the GPDD \citep[e.g.][]{connors2014} and the chosen priors in previous analyses \citep[e.g.][]{thorson2014a} and to greatly speed up chain convergence when running our model across all populations, we placed a weakly informative prior on $\phi$ that assumed the greatest probability density near zero with the reduced possibility of $\phi$ being near $-1$ or $1$. Specifically, we chose $\phi \sim \mathrm{Truncated\mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. The MCMC chains did not converge for \modelsNoConvergeAROne\ populations according to our criteria ($\widehat{R} < 1.05, n_\mathrm{eff} > 200$) after 8000 iterations of four chains. This included only \modelsNoConvergeAROneHeavyBase\ populations in which Pr($\nu < 10$) $> 0.5$ categorized them as heavy in the main Gompertz model. We did not include these models in Fig.~\ref{fig:alt}. \textit{TODO perhaps I should go up to 16,000 so that more converge.}

@@

We also fit a Ricker-logistic model:
\begin{align*}
x_t &= x_{t-1} + r_{\mathrm{max}}\left(1 - \frac{N_{t-1}}{K}\right) + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma),
\end{align*}
where $r_\mathrm{max}$ represents the theoretical maximum population growth rate that is obtained when $N_t$ (abundance at time $t$) $= 0$. The parameter $K$ represents the carrying capacity and, as before, $x_t$ represents the $\log$ transformed abundance at time $t$. The Ricker-logistic model assumes a linear decrease in population growth rate with increases in abundance. In contrast, the Gompertz model assumes a linear decrease in population growth rate with increases in \textit{log} abundance ($x_t$) \citep[e.g.][]{thibaut2012}.

@@

First, we tested the ability to estimate $\nu$ given different true values of $\nu$ and different sample sizes. We took stochastic draws from t distributions with different $\nu$ values ($\nu = 3, 5, 10,$ and $10^6$ [$\approx$ normal]), with central tendency parameters of $0$, and scale parameters of $1$. We started with $1600$ stochastic draws and then fit the models again at the first $800, 400, 200, 100, 50,$ and $25$ draws. Each time we recorded the posterior samples of $\nu$.

@@

We then fit our Gompertz models to the simulated datasets with all parameters (except $\nu$) set near the median values estimated in the GPDD. We repeated this with $50$ and $100$ samples without observation error, $50$ samples with observation error ($\sigma_\mathrm{obs} = 0.2$), and $50$ samples with the same observation error and a Gompertz model that allowed for correctly specified observation error magnitude.

@@

We fit a multilevel beta regression model to the predicted probability of heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed dynamics. The beta distribution is useful when response data range on a continuous scale between zero and one. We used a logit link function as is typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$ represents the predicted value for population $i$, class $j$, order $k$, and species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$ represents a vector of predictors (such as lifespan) for population $i$ with associated $\beta$ coefficients. The intercepts are allowed to vary from the overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$), taxonomic order ($\alpha^\mathrm{order}_k$), and species ($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \; \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \; \mathrm{species}}$. Where possible, we also allowed for error distributions around the predictors by incorporating the standard deviation of the posterior samples for the Gompertz parameters $\lambda$, $b$, and $\log \sigma$ around the mean point value as normal distributions (not shown in the above equation).

@@

We incorporated weakly informative priors into our model: $\mathrm{Cauchy}(0,
10)$ on the global intercept $\alpha$, $\mathrm{Half\mhyphen Cauchy}(0, 2.5)$ on all standard deviation parameters, $\mathrm{Half\mhyphen Cauchy}(0, 10)$ on the dispersion parameter $\phi_\mathrm{disp}$, and $\mathrm{Cauchy}(0, 2.5)$ on all other parameters \citep{gelman2006c, gelman2008d}. Compared to normal priors, the Cauchy priors concentrate more probability density around expected parameter values while allowing for a higher probability density far into the tails, thereby allowing the data to dominate the posterior more strongly if it disagrees with the prior. Our conclusions were not qualitatively changed by using uniform priors. We fit our models with 5000 total iterations per chain, 2500 warm-up iterations, four chains, and discarding every second sample to save memory. We checked for chain convergence visually and with the same criteria as before ($\widehat{R} < 1.05$ and $n_\mathrm{eff} >200$ for all parameters).

To derive taxonomic-order-level estimates of the probability of heavy tails accounting for time-series length (Fig \ref{fig:order-estimates}), we fit a separate multilevel model with the same structure but with only $\log$ time-series length as a predictor. (In this case, we did not want to control for intrinsic population characteristics such as density dependence.) Since our predictors were centered by subtracting their mean value, we obtained order-level estimates of the probability of heavy tails at mean log time-series length by adding the posteriors for $\alpha$, $\alpha^\mathrm{class}_j$, and $\alpha^\mathrm{order}_k$.
